---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

##Introduction
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

The variables included in this dataset are:
Dataset:  https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

The variables included in this dataset are:
steps: Number of steps taking in a 5-minute interval (missing values are coded as 𝙽𝙰)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

This report will finish the following tasks:
1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


##load the necessary packages for enviorment

```{r echo = TRUE}
library(dplyr)
library(knitr)
library(ggplot2)
library(lattice)
```
##Code for reading in the dataset and/or processing the data
If you need download this dataset, you can delte all the "#" before. Here I read
the data directly from the "activity.csv" in my workdirectory.

```{r echo = TRUE}
#fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(fileurl,destfile = "./act.zip", method = "curl")
#unlink(fileurl)
#unzip(zipfile = "./act.zip")
actData <- read.csv("activity.csv",header = TRUE,colClasses=c("numeric", "character", "numeric"))
```
Tidy the data classess
```{r echo=TRUE}
actData$date <- as.Date(actData$date,"%Y-%m-%d")
actData$interval<- as.factor(actData$interval)
```
##Histogram of the total number of steps taken each day
For this part of the assignment, we ignore the missing values in the dataset.
we use aggregate function to calculate the sum of steps taken each day and then we use 
ggplot function to draw the histogram. Of course, you should first install ggplot2 packages and library () it, if you does not own one.
```{r echo=TRUE}
stepsperday <- aggregate(actData$steps, by= list(actData$date),sum,na.rm = TRUE)

names(stepsperday) <- c("date","steps")

head(stepsperday)

month <- as.POSIXlt(stepsperday$date)$mon+1

day <- as.POSIXlt(stepsperday$date)$mday

histplot <- ggplot(data= stepsperday,aes(x = factor(day),fill = factor(month), y= steps)) + 
        geom_bar(stat = "identity") +
        labs(title = "Total Number of Steps per Day") + 
        labs(x= "Days",y="Total Numbers per Day") + 
        scale_fill_discrete(name = "Month")

print(histplot)
```

##Mean and median number of steps taken each day

```{r echo=TRUE}

meanstepsperday <- aggregate(actData$steps, by = list(actData$date),mean,na.rm = TRUE)

names(meanstepsperday) <- c("date","mean")

actData1 <- actData[-which(actData$steps == 0),]

medianstepsperday <- aggregate(actData1$steps, by = list(actData1$date),FUN = median,na.rm = TRUE)

names(medianstepsperday) <- c("date","median")

MMstepsperday <- data.frame(Mean = meanstepsperday$mean,Median = medianstepsperday$median, Date = meanstepsperday$date)

mean1 <- mean(actData$steps,na.rm = TRUE)

median1 <- median(actData1$steps,na.rm = TRUE)

print(MMstepsperday)

print(mean1)

print(median1)

```
##Time series plot of the average number of steps taken

Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Aggregate the steps per interval (i.e. 05, 10, 15…) and also calculates the mean for each interval. This calculated mean value is then plotted as a time series.

```{r echo=TRUE}

intervalSteps <- aggregate(actData$steps,by = list(actData$interval),mean, na.rm = TRUE)

names(intervalSteps) <- c("interval", "steps")

intervalSteps$interval <- as.numeric(intervalSteps$interval)

timeplot <- ggplot(intervalSteps,aes(x = interval,y = steps)) +
        geom_line(color = "orange", size = 1) + 
        labs(x="5-minute interval",y= "steps",title="Average Number of Steps Taken")

print(timeplot)

```

##The 5-minute interval that, on average, contains the maximum number of steps
There are a number of days/intervals where there are missing values (coded as 𝙽𝙰). The presence of missing days may introduce bias into some calculations or summaries of the data.

First, we calculate and report the total number of missing values in the dataset (i.e. the total number of rows with 𝙽𝙰s)

```{r echo= TRUE}
maxInterval <- intervalSteps[which.max(intervalSteps$steps),]

print(maxInterval)

```

##Code to describe and show a strategy for imputing missing data

Here,we use the mean/median for that day(yo can try other methods, such as median for that day or 5-minitue interval mean .ect.).Create a new dataset named Mergeddata that is equal to the original dataset but with the missing data filled in

```{r echo=TRUE}
newactData <- actData

newmeansteps <- aggregate(newactData$steps, by = list(newactData$interval),mean, na.rm= TRUE)

names(newmeansteps) <- c("interval","steps")

Mergeddata <- merge(newactData,newmeansteps,by = "interval")

Mergeddata$steps.x[is.na(Mergeddata$steps.x)] <- Mergeddata$steps.y[is.na(Mergeddata$steps.x)]

names(Mergeddata) <- c("interval","steps","date","meanstep")

Mergeddata1 <- Mergeddata[-which(Mergeddata$steps == 0),]

mean2 <- mean(Mergeddata$steps, na.rm=TRUE)

median2 <- median(Mergeddata1$steps,na.rm=TRUE)

print(mean2)

print(median2)

```
Here, we can compare the mean and median with first part. we can find that mean1 is equal to mean2 because that we omit all the NA in both calculate. However, the median2 is smaller than median1 because that we fill all NA's with mean.

##Histogram of the total number of steps taken each day after missing values are imputed
Here, we plot the histogram for the total number of steps taken each day after missing values are imputed:

```{r echo=TRUE}
stepsfilled <- aggregate(Mergeddata$steps,by = list(Mergeddata$date),sum,na.rm=TRUE)

names(stepsfilled) <- c("datef","stepsf")

head(stepsfilled)

monthf <- as.POSIXlt(stepsfilled$datef)$mon+1

dayf <- as.POSIXlt(stepsfilled$datef)$mday


histplotfilled <- ggplot(data= stepsfilled,aes(x = factor(dayf),fill = factor(monthf), y= stepsf)) + 
        geom_bar(stat = "identity") +
        labs(title = "Total Number of Steps per Day After Filled") + 
        labs(x= "Days",y="Total Numbers per Day") + 
        scale_fill_discrete(name = "Month")

print(histplotfilled)
```

##Panel plot comparing the average number of steps taken per 5-minute interval across 
##weekdays and weekends

Use the dataset with the filled-in missing values for this part.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.Make a panel plot containing a time series plot (i.e. 𝚝𝚢𝚙𝚎 = "𝚕") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r echo=TRUE}
wdays <- as.POSIXlt(actData$date)$wday

wdaysclass <- rep(0, length(actData$date))

wdaysclass[wdays>=1 & wdays <= 5] <- 1 #weekdays

wdaysclass[wdays ==6 | wdays == 0] <-2 #weekend

dayfactor <- factor(wdaysclass,levels = c(1,2),labels = c("weekdays","weekends"))

newactData$days <- dayfactor

WWstepsmean <- aggregate(steps~interval+days,newactData,mean, na.rm= TRUE)

xyplot(steps ~ interval | factor(days), data = WWstepsmean,aspect = 1/2, type = "l",col = "blue")
```
